Optimal (Multiway) Spatial Joins

### 2024-06-02 msj.tex
lmm 5.1

\noindent {\bf Proof of Lemma~\ref{lmm:hv:type2:recur-output}.} We will prove the lemma by mapping each tuple of $\J(R_1,...,R_{k-3},H',V)$ to a unique tuple in $\J_2$.

To describe the mapping, consider an arbitrary $(k-1)$-tuple $\bm{t} = (r_1,...,$ $r_{k-3},$ $h', v) \in \J(R_1,...,$ $R_{k-3},H',V)$. Let $h$ be the full segment of $h'$. By the definition of a trimmed segment (see Section~\ref{sec:hv:type2}), at least one rectangle $r_{k-2} \in R_{k-2}$ covers both the left endpoint of $h$ and the right endpoint of $h'$. We map $\bm{t}$ to the $k$-tuple $(r_1,...,r_{k-3}, r_{k-2},h,v)$. It is worth noting that $r_{k-2}$ definitely contains the entire $h'$ (which has the same left point as $h$, implying that $r_{k-2}$ covers both endpoints of $h'$).

We argue that $(r_1,...,r_{k-3}, r_{k-2},h,v)$ belongs to $\J_2$. Let $p$ be $h' \cap v$ (the intersection point of $h'$ and $v$). As $h'$ is contained in $h$, it must hold that $h \cap v = p$. Since $(r_1,...,r_{k-3}, h', v) \in \J(R_1,...,R_{k-3},H',V)$, we know that $p$ falls in all of $r_1, ..., r_{k-3}$. Our earlier discussion already shows that $r_{k-2}$ contains $h'$; this means that $p$ falls in $r_{k-2}$. Thus, $h \cap v \cap \bigcap_{i = 1}^{k-2}r_i$ cannot be empty as it equals $p$. We can now conclude that $(r_1,...,r_{k-3}, r_{k-2},h,v) \in \J(R_1,...,R_{k-2},H,V)$. This, together with the fact that $r_{k-2}$ covers the left endpoint of $h$, tells us $(r_1,...,r_{k-3}, r_{k-2},h,v) \in \J_2$.

It remains to show that no two distinct $(k-1)$-tuples $\bm{t}_1$ and $\bm{t}_2$ can be mapped to the same tuple in $\J_2$. Assume, on the contrary, that $\bm{t}_1$ and $\bm{t}_2$ are mapped to $(r_1,...,r_{k-3}, r_{k-2},h,v) \in \J_2$. Then, $\bm{t}_1[i] = \bm{t}_2[i] = r_i$ for each $i \in [k-3]$, and $\bm{t}_1[k-1] = \bm{t}_2[k-1] = v$. Moreover, $\bm{t}_1[k-2]$ and $\bm{t}_2[k-2]$ are both the trimmed segment of $h$. But then this means $\bm{t}_1 = \bm{t}_2$, giving a contradiction.

We can now conclude that $|\J(R_1,...,R_{k-3},H',V)| \le |\J_2| \le \out$, where the second inequality used the definition in \eqref{eqn:hv:type2:J2}.

@@

>>>>

We first prove Lemma~\ref{lmm:hv:type2:properties} and then proceed to prove Lemma~\ref{lmm:hv:type2:recur-output}, because the latter proof requires results in the former.

% \noindent {\bf Proof of Lemma~\ref{lmm:hv:type2:recur-output}.} We will prove the lemma by mapping each tuple of $\J(R_1,...,R_{k-3},H',V)$ to a unique tuple in $\J_2$.

% To describe the mapping, consider an arbitrary $(k-1)$-tuple $\bm{t} = (r_1,...,$ $r_{k-3},$ $h', v) \in \J(R_1,...,$ $R_{k-3},H',V)$. Let $h$ be the full segment of $h'$. By the definition of a trimmed segment (see Section~\ref{sec:hv:type2}), at least one rectangle $r_{k-2} \in R_{k-2}$ covers both the left endpoint of $h$ and the right endpoint of $h'$. We map $\bm{t}$ to the $k$-tuple $(r_1,...,r_{k-3}, r_{k-2},h,v)$. It is worth noting that $r_{k-2}$ definitely contains the entire $h'$ (which has the same left point as $h$, implying that $r_{k-2}$ covers both endpoints of $h'$).

% We argue that $(r_1,...,r_{k-3}, r_{k-2},h,v)$ belongs to $\J_2$. Let $p$ be $h' \cap v$ (the intersection point of $h'$ and $v$). As $h'$ is contained in $h$, it must hold that $h \cap v = p$. Since $(r_1,...,r_{k-3}, h', v) \in \J(R_1,...,R_{k-3},H',V)$, we know that $p$ falls in all of $r_1, ..., r_{k-3}$. Our earlier discussion already shows that $r_{k-2}$ contains $h'$; this means that $p$ falls in $r_{k-2}$. Thus, $h \cap v \cap \bigcap_{i = 1}^{k-2}r_i$ cannot be empty as it equals $p$. We can now conclude that $(r_1,...,r_{k-3}, r_{k-2},h,v) \in \J(R_1,...,R_{k-2},H,V)$. This, together with the fact that $r_{k-2}$ covers the left endpoint of $h$, tells us $(r_1,...,r_{k-3}, r_{k-2},h,v) \in \J_2$.

% It remains to show that no two distinct $(k-1)$-tuples $\bm{t}_1$ and $\bm{t}_2$ can be mapped to the same tuple in $\J_2$. Assume, on the contrary, that $\bm{t}_1$ and $\bm{t}_2$ are mapped to $(r_1,...,r_{k-3}, r_{k-2},h,v) \in \J_2$. Then, $\bm{t}_1[i] = \bm{t}_2[i] = r_i$ for each $i \in [k-3]$, and $\bm{t}_1[k-1] = \bm{t}_2[k-1] = v$. Moreover, $\bm{t}_1[k-2]$ and $\bm{t}_2[k-2]$ are both the trimmed segment of $h$. But then this means $\bm{t}_1 = \bm{t}_2$, giving a contradiction.

% We can now conclude that $|\J(R_1,...,R_{k-3},H',V)| \le |\J_2| \le \out$, where the second inequality used the definition in \eqref{eqn:hv:type2:J2}.

@@

\extraspacing \noindent {\bf Proof of Lemma~\ref{lmm:hv:type2:recur-output}.}
First, we prove that $\cross_{R_{k-2}}(\bm{t})$ is non-empty for each tuple $\bm{t} \in \J(R_1,...,R_{k-3},H',V)$. Let $\bm{t} = (r_1,...,r_{k-3},h',v)$. By the definition of a trimmed segment, there exists a rectangle $r_{k-2} \in R_{k-2}$ that covers both the left endpoint of $h$ and the right endpoint of $h'$. Thus, $r_{k-2}$ contains $h'$, because $h'$ shares the same left endpoint as $h$. Since the effective horizontal segment of $\bm{t}$ is contained in $h'$, it is also covered by $r_{k-2}$. This implies that $\cross_{R_{k-2}}(\bm{t})$ includes $r_{k-2}$, and is therefore non-empty.

Then, we resorts to the algorithm generate-$\J_2$ (see Section~\ref{sec:hv:type2}). For each tuple $\bm{t} \in \J(R_1,...,R_{k-3},H',V)$, this algorithm adds to $\J_2$ exactly $|\cross_{R_{k-2}}(\bm{t})|\geq 1$ tuples. Therefore, $|\J(R_1, ..., R_{k-3}, H', V)| \le |\J_2| \le \out$, where the last inequality follows from the definition of $\J_2$ (see \eqref{eqn:hv:type2:J2}).
